---
title: "PA_new"
output: html_document
date: "2025-03-11"
---

## Protected area boundary change

This notebook outlines the methodology used to measure at protected area boundaries via GEE. 

The notebook queries MODIS imagery and returns the gradient values of relevant bands as well as the vegetation indicies NDVI and EVI. This code expects an annual time span and returns a geoTIFF for each band in each park for that year. This notebook is designed to test the code, which then is scaled to retrieve the full annual timeseries. 

```{r}
library(tidyverse)
library(rgee)
```

```{r}
ee_Authenticate()
ee_Initialize(project = 'dse-staff')
print(ee$String('Hello from the Earth Engine servers!')$getInfo())
```

#### Load dataset 
```{r}
# Load a FeatureCollection (example: World Database on Protected Areas)
protected_areas <- ee$FeatureCollection("WCMC/WDPA/202106/polygons")

# Define individual filters
marine_filter <- , "0")ee$Filter$eq("MARINE" # Remove marine areas
not_mpa_filter <- ee$Filter$neq("DESIG_ENG", "Marine Protected Area")  # Exclude MPAs
status_filter <- ee$Filter$inList("STATUS", c("Designated", "Established", "Inscribed")) 
designation_filter <- ee$Filter$neq("DESIG_ENG", "UNESCO-MAB Biosphere Reserve")
excluded_pids <- c("555655917", "555656005", "555656013", "555665477", "555656021",
                   "555665485", "555556142", "187", "555703455", "555563456", "15894") 
area_filter <- ee$Filter$gte("GIS_AREA", 200) # larger than 200 km^2

# Filter to exclude specific WDPA_PID values
pids_filter <- ee$Filter$inList("WDPA_PID", excluded_pids)$Not()

# Combine filters using logical operations
combined_filter <- ee$Filter$And(
  marine_filter,
  not_mpa_filter,
  status_filter,
  designation_filter,
  pids_filter,
  area_filter
)

# Apply the combined filter to the FeatureCollection
data <- protected_areas$filter(combined_filter)

# Select the WDPA_PID property from the filtered dataset
wdpa_pids <- data$aggregate_array("WDPA_PID")  # Extracts all values of the WDPA_PID property as a list

# Retrieve the list as an R vector
wdpa_pid_list <- wdpa_pids$getInfo()

# Print the list of WDPA_PIDs
print(length(wdpa_pid_list)) #9272
```

#### Set global constants

```{r}
PROTECTED_AREAS <- data
WATER_MASK <- ee$Image('JRC/GSW1_0/GlobalSurfaceWater')
MODIS <- ee$ImageCollection('MODIS/006/MOD09A1')
BUFFER <- 10000
MAX_ERR <- 1
YEAR <- 2010
```

#### Define functions
```{r}
buffer_polygon <- function(feat, buff, err){
  feat <- ee$Feature(feat)
  out <- feat$buffer(buff)$geometry()  # 10km out
  inn <- feat$buffer(-buff)$geometry()  # 10km in
  aoi <- out$difference(inn, err)
  return(aoi)
}
```

```{r}
mask_water <- function(feat){
  water_no_holes <- WATER_MASK$select('max_extent')$
    focalMax(radius = 30, units = 'meters', kernelType = 'square')$
    focalMin(radius = 30, units = 'meters', kernelType = 'square')
  water_vect <- water_no_holes$reduceToVectors(
    reducer = ee$Reducer$countEvery(),
    geometry = feat$buffer(1000),  # Ensures water is removed at edges
    scale = 30,
    maxPixels = 1e10,
    geometryType = 'polygon',
    eightConnected = FALSE)
  geom <- feat$difference(right = water_vect$geometry(), maxError = 1)
  return(geom)
}
```

```{r}
filter_for_year <- function(feat, year) {
  start <- ee$Date$fromYMD(year, 1, 1)
  return(ee$Filter$And(
    ee$Filter$bounds(feat),
    ee$Filter$date(start, start$advance(1, "year"))
  ))
}
```

```{r}
add_indices_to_image <- function(image) {
  
  EVI <- image$expression(
    "2.5 * ((NIR - RED) / (NIR + 6 * RED - 7.5 * BLUE + 1))",
    list(
      NIR = image$select("sur_refl_b02"),  # MODIS NIR
      RED = image$select("sur_refl_b01"),  # MODIS RED
      BLUE = image$select("sur_refl_b03")  # MODIS BLUE
    )
  )$rename("EVI")
  
  NDVI <- image$expression(
    "(NIR - RED) / (NIR + RED)",
    list(
      NIR = image$select("sur_refl_b02"),
      RED = image$select("sur_refl_b01")
    )
  )$rename("NDVI")
  
  image <- image$addBands(list(EVI, NDVI))
  
  #print(paste("All band names:", image$bandNames()$getInfo()))
  
  return(image)
}
```

```{r}
get_gradient_magnitude <- function(image){
  gradient <- image$gradient()
  gradient_x <- gradient$select('x') 
  gradient_y <- gradient$select('y') 
  magnitude <- gradient_x$pow(2)$add(gradient_y$pow(2))$sqrt() 
  return(magnitude)
}
```

```{r}
geometry <- ee$Geometry$Polygon(list(
  list(
    c(-119.80738088038615, 38.53793770547406),
    c(-120.35120412257365, 38.64956590608609),
    c(-120.86206838038615, 38.51215270703278),
    c(-120.41162892726115, 38.05942162797243),
    c(-120.13697072413615, 38.24085287086409),
    c(-120.05457326319865, 37.98585693208372),
    c(-119.85681935694865, 38.01182937540273),
    c(-119.79639455226115, 38.11130517090483),
    c(-119.71399709132365, 38.119948848017394),
    c(-119.63709279444865, 38.1674707827119),
    c(-119.62061330226115, 38.2494811940204),
    c(-119.65357228663615, 38.32709000484806),
    c(-119.65906545069865, 38.38739502408541),
    c(-119.60962697413615, 38.46915718247208),
    c(-119.69751759913615, 38.49065815171371),
    c(-119.90076466944865, 38.3012296223024),
    c(-119.94470998194865, 38.3658632836034),
    c(-119.97217580226115, 38.41322465329759)
  )
))
```

```{r}
serengeti <- data$
  filter(ee$Filter$eq('ORIG_NAME', 'Serengeti National Park'))$
  first()
serengeti_geometry <- single_feature$geometry()
```

```{r}
serengeti_geometry <- serengeti$geometry()

# Now we can visualize the geometry using rgee
Map$centerObject(aoi, zoom = 8)
Map$addLayer(aoi, list(color = 'red'), 'Serengeti Geometry')
```


```{r}
aoi <- buffer_polygon(serengeti_geometry, BUFFER, MAX_ERR)|> mask_water() 
modis_ic <- MODIS$filter(filter_for_year(aoi, YEAR))
band_names <- modis_ic$first()$bandNames()
composite <- modis_ic$reduce(ee$Reducer$median())$rename(band_names)$clip(aoi)
image <- add_indices_to_image(composite)
single_band <- image$select("sur_refl_b01")
gradient <- get_gradient_magnitude(single_band)
# Normalize the gradient magnitude to the range 0-1
max_val <- gradient$reduce(ee$Reducer$max())  # Find the maximum value in the gradient image
normalized_gradient <- gradient$divide(max_val)  # Divide by the max value to normalize
```

```{r}
vis_params <- list(
  min = -0.5, 
  max = 1, 
  palette = c('blue', 'white', 'green')
)

# Add single_band (NDVI) to the map
Map$centerObject(aoi, zoom = 8)
Map$addLayer(single_band, vis_params, "NDVI Layer")
```


```{r}
# Step 3: Apply thresholding (simple example with a fixed threshold)
threshold_value <- 0.1
edges <- gradient$gt(threshold_value)

# Apply Gaussian blur to smooth the gradient magnitude
smoothed_gradient <- gradient$focal_mean(radius = 3, kernelType = "circle", units = "pixels")

# Step 4: Visualize the edges
vis_params <- list(
  min = 0,
  max = 1,
  palette = c('blue', 'white', 'red')
)

# Map the result using Leaflet
Map$centerObject(aoi, zoom = 10)
Map$addLayer(gradient, vis_params)
```

```{r}
vis_params <- list(
  min = 0,
  max = 1,
  palette = c('blue', 'white', 'red')
)

# Map the result using Leaflet
Map$centerObject(aoi, zoom = 10)
Map$addLayer(normalized_gradient, vis_params)
```

```{r}
download_task <- ee$batch$Export$image$toDrive(
  image = edges,
  description = 'Canny_Edges',
  folder = 'earth_engine_outputs',
  scale = 30,
  region = aoi
)
```

#### Visualization (OPTIONAL)

```{r}
export_image_to_cloud <- function(image) {
  
  export_task <- ee$batch$Export$image$toCloudStorage(
    image = image,
    description = "gradient",
    bucket = "dse-staff",
    fileNamePrefix = "gradient",
    fileFormat = "GeoTIFF",
    formatOptions = list(cloudOptimized = TRUE),
    maxPixels = 1e9,
    scale = 30
  )
  export_task$start()
  print("saved")
}

export_image_to_cloud(gradient)
```


#### Testing 

```{r}
line <- ee$Geometry$LineString(geometry$coordinates()$get(0))
line_buffered <- line$buffer(500)  # 500 meters buffer

pixel_median_value_line <- single_band$reduceRegions(
  collection = ee$FeatureCollection(line),
  reducer = ee$Reducer$median(),
  scale = 500 
)

pixel_median_value_buffer <- single_band$reduceRegions(
  collection = ee$FeatureCollection(aoi),
  reducer = ee$Reducer$median(),
  scale = 500 
)

pixel_count_line <- image$reduceRegion(
  reducer = ee$Reducer$count(),
  geometry = line_buffered,
  scale = 500,  # Match dataset resolution
  maxPixels = 1e13
)

pixel_count_buffer <- image$reduceRegion(
  reducer = ee$Reducer$count(),
  geometry = aoi,
  scale = 500,  # Match dataset resolution
  maxPixels = 1e13
)

# Print results
pixel_median_value_line$getInfo()
pixel_median_value_buffer$getInfo()
pixel_count_line$getInfo()
pixel_count_buffer$getInfo()
```


```{r}
# Load the required library
library(rgee)

# Initialize the Earth Engine API
ee_Initialize()

# Define the region and date range for MODIS data
region <- ee$Geometry$BBox(-122.5549, 37.6968, -122.3446, 37.8111)
start_date <- '2016-01-01'
end_date <- '2022-12-31'

# Load MODIS data (MOD09A1)
collection <- ee$ImageCollection("MODIS/006/MOD09A1")$
  filterDate(start_date, end_date)$
  filterBounds(region)

# Reduce the collection by taking the median
images <- collection$reduce(ee$Reducer$median())

# Define visualization parameters for RGB bands (assuming B1, B4, B3 correspond to Red, Green, and Blue)
vis_params <- list(
  min = 0, 
  max = 10000,  # Typical scaling for MODIS surface reflectance bands
  bands = c("sur_refl_b01_median", "sur_refl_b04", "sur_refl_b03")  # B1: Red, B4: Green, B3: Blue
)

# Map visualization
Map$centerObject(region)
Map$addLayer(images, vis_params, "MODIS RGB", FALSE)

# Show the map
Map

```


```{r}
get_polygon_annual_median_perpixel <- function(geom, start_date, end_date, show_map = FALSE) {
  
  # Load MODIS surface reflectance dataset
  ic <- ee$ImageCollection("MODIS/006/MOD09A1")$  
    filterBounds(geom)$
    filterDate(start_date, end_date)$
    select(c("sur_refl_b01", "sur_refl_b02", "sur_refl_b03","sur_refl_b04", 
             "sur_refl_b05", "sur_refl_b06", "sur_refl_b07"))  
  
  image <- ic$mean()$clip(geom)
  
  image <- add_indices_to_image(image)
  
  image <- get_gradient_magnitude(image)
  
  image <- image$set(
    list(
      "system:time_start" = ee$Date(start_date)$millis(),
      "system:time_end" = ee$Date(end_date)$millis()
    )
  )
  
  # Display image and polygon on map if requested
  if (show_map) {
    vizParams <- list(
      bands = c("sur_refl_b01", "sur_refl_b02", "sur_refl_b03"),  # RED, NIR, BLUE
      min = 0,
      max = 5000,  # MODIS surface reflectance scaling
      gamma = c(0.95, 1.1, 1)
    )
    
    Map$centerObject(geom, 6)
    Map$addLayer(image, vizParams, "MODIS False Color Composite")
    Map$addLayer(geom, list(color = "red"), "Polygon")
  }
  
  return(image)
}
```

